---
title: "Assignment 04"
author: "Hannah Hill & Imani Sherill"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```
[U.S. Census Bureau Income Inequality](https://www.census.gov/library/publications/2021/demo/p60-273.html)

```{r}
library(tidyverse)
library(srvyr)
library(janitor)
library(dplyr)
census <- read_csv("poverty1.csv")
census <- census [9:63 , 1:5] %>%
  na.omit() %>%
  rename(category = 1) %>%
  rename(total_pop = 2) %>%
  rename(in_poverty = 3) %>%
  rename(moe = 4) %>%
  rename(percent_poverty = 5)
view(census)
```

```{r, cleaning data}
race <- census [2:6, 1:5] %>%
  rename(race = category)
age <- census [9:11, 1:5] %>%
  rename(age = category)
sex <- census [7:8, 1:5] %>%
  rename(sex = category)
nativity <- census [12:15, 1:5] %>%
  rename(nativity = category)
region <- census [16:19, 1:5] %>%
  rename(region = category)
metro <- census [20:23, 1:5] %>%
  rename(metro = category)
work_status <- census [25:28, 1:5] %>%
  rename(work_status = category)
disability <- census [30:31, 1:5] %>%
  rename(disability = category)
educ <- census [33:36, 1:5] %>%
  rename(educ = category)
```


```{r}
gini <- read_csv("gini.csv")
gini <- gini [4:63, 1:14] %>%
  rename(category = 1)
gini <- gini [4:63, 1:14] %>%
  na.omit() %>%
  rename(year = category) %>%
  rename(mean_lowest_quint = 2) %>%
  rename(mean_second_quint = 3) %>%
  rename(mean_middle_quint = 4) %>%
  rename(mean_fourth_quint = 5) %>%
  rename(mean_highest_quint = 6) %>%
  rename(mean_top5 = 7) %>%
  rename(gini_coefficient = 14)
giniplot <- gini [1:56, 1:7]

inequality <- read_csv("inequality.csv")
#probably don't use inequality just use gini and census
```

# Measuring Income Inequality in the United States, 1976-2020
## Data Visualization 01

The Gini coefficient is a measure of income inequality across different groups. A Gini coefficient closer to 0 represents equal income while a value closer to 1 represents a high degree of inequality. In the graph below, income inequality steadily increases from .38 to to .47, representing a shift in wider dispersion and greater inequality between 1967 and 2020 within the United States. 

``` {r}
library(ggplot2)
gini %>%
  mutate(gini_coefficient = as.numeric(gini_coefficient)) %>%
  mutate(year = as.numeric(year)) %>%
  ggplot(mapping = aes(x = year, y = gini_coefficient)) +
  geom_point(color = "blue", size=3, shape=18) +
  geom_smooth(method=lm, color = "red", size=.2, se=FALSE) +
  ylim(.25,.55) +
  theme_minimal() +
  labs(
    title = "Income Inequality in the United States",
    subtitle = "1967-2020",
    caption = "U.S. Census Bureau",
    x = "Year",
    y = "Gini Coefficients") +
  theme_minimal()
```

# The Rich Keep Getting Richer
## Data Visualization 02

In Figure 2, the mean income for the top 5% of households in the United States is documented across the years 1967 to 2020. While these data provide a small insight into how the wealthiest are doing financially the last 50 years, analysis is limited by the fact that the data used are averages and not medians. This graph, however, does illustrate the impact of the 2008 Great Recession had on even the top 5% of households where mean household income drops off sharply after 2007. Despite being impacted by the Great Recession, overall growth for the mean household income trends upwards with a **133.5% increase** between 1967 and 2020. In contrast, Figure 5 shows the bottom quintile mean household income over the same time period. Here, the impact of the 2008 Great Recession is much more noticeable for low income households where income remains stagnant for several years. For these households, mean household income has only **increased by 34.2%** between 1967 and 2020.
```{r}
giniplot %>%
  mutate(year = as.numeric(year)) %>%
  mutate(mean_top5 = str_replace_all(mean_top5, ",", "")) %>%
  mutate(mean_top5 = as.numeric(mean_top5)) %>%
  ggplot(aes(x = year, y = mean_top5)) +
  geom_area(fill = "steelblue", alpha =0.4) +
  geom_line(color = "steelblue", size = 2) +
  geom_point(size = 1.5, color = "blue") +
  geom_jitter() +
  scale_x_continuous(breaks = c(1975,1990,2005,2020)) +
  scale_y_continuous(breaks = c(0,55000,120000,185000,250000, 315000, 380000,445000)) +
  theme_minimal() +
  labs(
    title = "Figure 2 - The Rich Keep Getting Richer:",
    subtitle = "The Top 5% of Household's Mean Income in the U.S., 1967-2020",
    caption = "U.S. Census Bureau",
    x = "Year",
    y = "Mean Annual Income"
  )
```

```{r}
giniplot %>%
  mutate(year = as.numeric(year)) %>%
  mutate(mean_lowest_quint = str_replace_all(mean_lowest_quint, ",", "")) %>%
  mutate(mean_lowest_quint = as.numeric(mean_lowest_quint)) %>%
  ggplot(aes(x = year, y = mean_lowest_quint)) +
  geom_area(fill = "maroon", alpha =0.4) +
  geom_line(color = "maroon", size = 2) +
  geom_point(size = 1.5, color = "red") +
  geom_jitter() +
  scale_x_continuous(breaks = c(1975,1990,2005,2020)) +
  theme_minimal() +
  labs(
    title = "Figure 5 - Stagnant Growth for the Lowest Earning Households:",
    subtitle = "The Bottom Quintile of Household's Mean Income in the U.S., 1967-2020",
    caption = "U.S. Census Bureau",
    x = "Year",
    y = "Mean Annual Income"
  )
```

```{r}
library(tidyverse)
library (haven)
library(janitor)
library(srvyr)
library (dplyr)
```

```{r}
##census <- read_excel(path = "relative/file/path/to/data")
census_poverty <- read_excel("poverty.xlsx") %>%
janitor::clean_names()
glimpse (census_poverty) 
```

```{r}
#data cleaning 
age <- census_poverty [ 1: 3, 1 : 5,  ] %>% 
  rename( age = category )
view (age)
nativity <- census_poverty [ 4 : 7, 1 : 5] %>% 
  rename (nativity = category)
view (nativity)
region <-  census_poverty [ 8 : 11, 1 :5 ] %>% 
  rename (region = category )
view (region )
metropolitan <-  census_poverty [12 : 15, 1 : 5] %>% 
  rename (metropolitan = category)
view (metropolitan)
employment <- census_poverty [17:21 , 1:5] %>% 
  rename (employment = category )
view (employment)
disability <-  census_poverty [ 22: 23, 1: 5] %>% 
  rename (disability = category)
view (disability)
educational_attainment <- census_poverty [25: 28, 1:5 ] %>% 
  rename (educational_attainment = category) 
view (educational_attainment)
```

```{r}
#transforming categorical variables to numeric 
age <- factor(c("under age 18", "aged 18 to 64", "aged 65 and older")) 
age 
age_num <- unclass(age)
age_num
view (nativity)
nativity <-  factor(c("native-born", "foreign-born", "naturalized citizen", "not a citizen"))
nativity_num <- unclass(nativity) 
nativity_num
view(region)
region <-  factor(c("Northeast", "Midwest", "South", "West"))
region_num <- unclass(region)
nativity_num
view (metropolitan)
metropolitan <-  factor(c("urban", "rural"))
metropolitan_num <- unclass(metropolitan)
metropolitan_num
view (employment)
employment <-  factor(c("all workers, worked full time", "part-time"))
employment_num <- unclass(employment)
employment_num
view (educational_attainment)
educational_attainment <- factor(c("no highschool diploma", "only highschool", "some college", "bachelor's degree or higher")) 
educational_attainment_num <- unclass(educational_attainment)
educational_attainment_num
```

# The Relationship Between Educational Attainment & Poverty
## Data Visualization 03

Figure 3 highlights the relationship between education and poverty in the United States in the year 2019. As educational attainment increases, the population's poverty percentage decreases. A highly discussed topic, education is now recognized as an important social and economic policy tool. Using education as a resource can help decrease poverty as shown in this graph. However, policymakers must be wary of stating that all educational attainment provides positive economic returns. Specifically, there is a stark difference in level of poverty between "no high school diploma" and "bachelor's degree". This graph does exemplify some diminishing marginal returns between "some college" and a "bachelor's degree or higher", but it emphasizes the importance of a bachelor's degree and its relevance to decreasing poverty within higher-education

```{r}

install.packages(ggrepel)
library(ggrepel)

educational_attainment %>%
  select(category, total_pop, in_poverty, percent_poverty) %>%
  mutate(percent_poverty = as.numeric(percent_poverty)) %>%
  ggplot() +
  geom_col(mapping = aes(x = category, y = percent_poverty, fill = category)) +
  theme_minimal() +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  labs(
    title = "The Relationship Between Educational Attainment & Poverty",
    subtitle = "The Importance of Higher Educational Attainment in Mitigating Poverty, 2019 ",
    caption = "U.S. Census Bureau 2019",
    x = "Educational Attainment",
    y = "Percentage_Poverty"
  )
```

#Poverty by Region in the United States 
## Data Visualization 04

Figure 4 shows percent poverty based off of regions in the United States in 2019. Historically, the South has the highest percentages of poverty in the United States. This graph supports those findings in 2019. The Northeast, Midwest, and West have statistically consistent poverty rates between the three regions. The difference in poverty by region lies in inaccessible resources. Southern states have higher poverty rates, as seen in this graph, lower educational attainment, and lack of economic mobility. Less metropolitan regions lie in the South creating less economic diversity and low median incomes.

```{r}

region %>% 
  mutate(percent_poverty = as.numeric(percent_poverty)) %>% 
ggplot(mapping=(aes(x="", y = percent_poverty, fill = category))) + 
          geom_bar(stat = "identity", width = 1, color = "white") + 
          #coord_polar(theta = "y")
          coord_polar ("y", start = 0) + 
  theme_void() + 
  #geom_text(aes(y = category, label = percent_poverty ) , color ="white", size = 4) +
  geom_text(aes(label = percent_poverty), 
            position = position_stack(vjust = 0.5)) +
  labs(
    title = "Poverty by Region in the United States",
    subtitle = "Southern Regions of the United States Continue to Hold High Poverty Rates",
    caption = "U.S. Census Bureau 2019",
  )
```

